---
title: "Thou shalt know thy data"
author: "Your name here"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---

The point of this exercise is to convince you that you should examine your data closely and know what it looks like before stuffing it into a fancy regression model. 

```{r include = FALSE}
rm(list = ls())
#library(xlsx)
library(tidyverse)
library(stargazer)
library(knitr)
library(pander)
#library(data.table)
#library(RCurl)

data <- read.csv("anscombe.csv")
                   
```


1. Create a table of summary statistics of the variables using stargazer: 
```{r echo = FALSE, message = FALSE, warnings = FALSE, results = "asis"}
###   Check out summary stats of the data


```


2. Create a set of scatter plots for each set of variables, by plotting y1 against x1, y2 against x2, etc. How do they look to you? Does it look like a linear regression model will be a good fit for each dataset?  
```{r echo = FALSE, message = FALSE, warnings = FALSE, results = "asis"}
###  These can be done separately, or you can transform the data to do a faceted plot


```


3. Now, run regressions on these sets. Collect the results into one stargazer table. What do you notice about the results? 

```{r echo = FALSE, message = FALSE, warnings = FALSE, results = "asis"}
###  if we want the numbers:


```




4. What would you conclude about trusting regression results without knowing what the data looks like first? 

